---
title: "An Exploratory Data and Network Analysis of Movies"
author: Clarice, Daven, Lucia, Christopher and Indurain
date: September 6, 2019
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{css, echo=FALSE}
code {
  font-size:0.8em;
}

table {
  font-size:0.9em;
}

p {
  font-size:0.9em;
}
```

## Introduction

In this report, we will be analysing a dataset from [Kaggle](https://www.kaggle.com/rounakbanik/the-movies-dataset#links.csv), which contains movies of different genres produced over a vast number of years. What makes this analysis interesting is that we can try and draw various conclusions based on a movie's popularity, directors or actors involved, year of production, and so forth. Moreover, we can construct various networks in an attempt to find meaningful and interesting results.
When inspecting a database of films from recent years, various interesting inferences are uncovered. A film may have a high rating yet low return on investment. Which genre would you guess is the most successful? Which actors do you think are the most popular? 


We have split our Exploratory Data Analysis into four main parts:

|  Section 	|                                                                                                                                                                                                                                                 	|
|---	|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------	|
| 1 	| Introducing the Data <br> - We first try to understand the data and look at its content.  |
| 2 	| Pre-Processing <br>- We look at what needs to be altered or removed from the dataset. <br>- We try clean any dirty text. <br>- We try to minimise the dataset's missing values.                                                                                    	|
| 3 	| Exploring the Data <br>- We conduct basic analysis on the dataset.  <br>- We explore genres. <br>- We explore movie popularity. <br>- We look at profit, gross, and return of interests with movies. <br>- We conduct more advanced analysis on the dataset.               	|
| 4 	| Network Analysis <br>- We measure the network (centrality, degree distribution, number of components, average degree) <br>- We use network measures to highlight certain nodes (actors) and see which measures of an actor will increase ratings and budgets.  	|                                                                                                                                  |

-----

## Admin

Before we start, let's keep this code chunk for importing the correct libraries and loading the appropriate dataset. We use pacman to load the following:

```{r}
pacman::p_load(rjson, jsonlite, DT,  RJSONIO, data.table, dplyr, compareDF, prettydoc, rmdformats, VIM, ggplot2, stringr, tidyr, plotly, RColorBrewer, formattable, corrplot)
```

We import the dataset like this:
```{r}
movie_metadata <- read.csv("../data/movie_metadata.csv", sep=";")
```

In the next section we can introduce our dataset and look its content. 

-----

## Introducing The Dataset

This section of the report is quite essential for our analysis. We cannot make any interesting inferences from the dataset if we do not know what is contained within it. In this section we will try to understand exactly what we are dealing with. Thereafter, we can begin to draw interesting results. We have already read in our dataset called `movie_metadata`, so we can see the following:

The dataset contains 28 unique columns/variables, each of which are described in the table below:

Variable Name            | Description
-------------------------|----------------------------------------------
color                    | Specifies whether a movie is in black and white or color
director_name            | Contains name of the director of a movie
num_critic_for_reviews   | Contains number of critic reviews per movie
duration                 | Contains duration of a movie in minutes
director_facebook_likes  | Contains number of facebook likes for a director
actor_3_facebook_likes   | Contains number of facebook likes for actor 3
actor_2_name             | Contains name of 2nd leading actor of a movie
actor_1_facebook_likes   | Contains number of facebook likes for actor 1 
gross                    | Contains the amount a movie grossed in USD
genres                   | Contains the sub-genres to which a movie belongs
actor_1_name             | Contains name of the actor in lead role
movie_title              | Title of the Movie
num_voted_users          | Contains number of users votes for a movie
cast_total_facebook_likes| Contains number of facebook likes for the entire cast of a movie
actor_3_name             | Contains the name of the 3rd leading actor of a movie
facenumber_in_poster     | Contains number of actors faces on a movie poster
plot_keywords            | Contains key plot words associated with a movie
movie_imdb_link          | Contains the link to the imdb movie page
num_user_for_reviews     | Contains the number of user generated reviews per movie
language                 | Contains the language of a movie
country                  | Contains the name of the country in which a movie was made
content_rating           | Contains maturity rating of a movie
budget                   | Contains the amount of money spent in production per movie
title_year               | Contains the year in which a film was released
actor_2_facebook_likes   | Contains number of facebook likes for actor 2
imdb_score               | Contains user generated rating per movie
aspect_ratio             | Contains the size of the aspect ratio of a movie
movie_facebook_likes     | Number of likes of the movie on its Facebook Page


Furthermore, the dataset contains `5043` movies, spanning accross `96` years in 46 countries. There are `1693` unique director names and `5390` number of actors/actresses. Around `79%` of the movies are from the USA, `8%` from UK, and `13%` from other countries.

The structure of the dataset can also be used to understand our data. We can run the following code chunk to see its structure.

```{r}
# Get structure of dataset
str(movie_metadata)
```

In the next section we can start preparing the dataset for analyis by removing or simplifying some of the data.

-----

## Pre-Processing Data

In this part of the report we attempt to look for various things that may have a negative or insignificant impact on the inferences we make on the dataset. Once we have sufficiently cleaned and prepared the dataset, we can commence with drawing various conclusions from the graphs we generate. 


### Duplicate Rows

In `movie_metadata`, we have some duplicate rows, so we want to remove the 45 duplicated rows and keep the unique ones. 

```{r}
# find duplicated rows
sum(duplicated(movie_metadata))
```

```{r Remove Duplicates}
# Remove duplicated rows
movie_metadata <- movie_metadata[!duplicated(movie_metadata), ]
```

### Missing Values

Let's have a look at the number of NA values in our dataset:
```{r}
# Find NA values
colSums(sapply(movie_metadata, is.na))
```

To help visualise this, have a look at the following heatmap of the missing values:
```{r fig.align="center"}
# Visualise Missing Values
missing.values <- aggr(movie_metadata, sortVars = T, prop = T, sortCombs = T, cex.lab = 1.5, cex.axis = .6, cex.numbers = 5, combined = F, gap = -.2)
```

#### Gross and Budget

Since `gross` and `budget` have too many missing values (874 and 488), and we want to keep these two variables for the following analysis, we can only delete rows with null values for gross and budget because imputation will not do a good job here.

```{r}
# Find NA values for gross and budget
movie_metadata <- movie_metadata[!is.na(movie_metadata$gross), ]
movie_metadata <- movie_metadata[!is.na(movie_metadata$budget), ]
dim(movie_metadata)
```

The difference in observations have decreased by `4998 - 3857 = 1141` which is luckily only `22.8%` of the previous total observations.
Let’s have a look at how many complete cases we have.

#### Content Rating

```{r}
# Look at all the different types of content ratings
table(movie_metadata$content_rating)
```

According to the history of naming these different content ratings, we find `M = GP = PG, X = NC-17`. We want to replace `M` and `GP` with `PG`, replace `X` with `NC-17`, because these two are what we use nowadays.

```{r}
movie_metadata$content_rating[movie_metadata$content_rating == 'M']   <- 'PG' 
movie_metadata$content_rating[movie_metadata$content_rating == 'GP']  <- 'PG' 
movie_metadata$content_rating[movie_metadata$content_rating == 'X']   <- 'NC-17'
```

We want to replace `Approved`, `Not Rated`, `Passed`, `Unrated` with the most common rating `R`.

```{r}
movie_metadata$content_rating[movie_metadata$content_rating == 'Approved']  <- 'R' 
movie_metadata$content_rating[movie_metadata$content_rating == 'Not Rated'] <- 'R' 
movie_metadata$content_rating[movie_metadata$content_rating == 'Passed']    <- 'R' 
movie_metadata$content_rating[movie_metadata$content_rating == 'Unrated']   <- 'R' 
movie_metadata$content_rating <- factor(movie_metadata$content_rating)
table(movie_metadata$content_rating)
```

Blanks should be taken as missing value. Since these missing values cannot be replaced with reasonable data, we delete these rows.

```{r}
# Remove rows with blank content ratings
movie_metadata <- movie_metadata[!(movie_metadata$content_rating %in% ""),]
```

### Delete (Some) Rows

```{r}
colSums(sapply(movie_metadata, is.na))
```

We remove `aspect_ratio` because 1 it has a lot of missing values and 2 we will not be looking into the impact that it has on other data (we assume that it doesn't). 

```{r}
# Remove aspect_ratio column
movie_metadata <- subset(movie_metadata, select = -c(aspect_ratio))
```

### Add a Column

#### Gross and Budget

We have gross and budget information. So let’s add two colums: profit and percentage return on investment for further analysis.

```{r}
# add profit and return of investment column
movie_metadata <- movie_metadata %>% 
  mutate(profit = gross - budget,
         return_on_investment_perc = (profit/budget)*100)
```

### Remove (Some) Columns

#### Colour

Next, we take a look at the influence of `colour` vs `black and white`. 

```{r}
# Get colour display types of movies
table(movie_metadata$color)
```

Since `3.4%`of the data is in black and white, we can remove the `color` column it. 

```{r}
# delete colour
movie_metadata <- subset(movie_metadata, select = -c(color))
```

#### Language

Let's have a look at the different languages contained within the dataset.

```{r}
# Look at different languages
table(movie_metadata$language)
```

Almost `95%` movies are in English, which means this variable is nearly constant. Let’s remove it.

#### Country

Next, we can look at the different types of countries.

```{r}
table(movie_metadata$country)
```

Around `79%` movies are from USA, `8%` from UK, `13%` from other countries. So we group other countries together to make this categorical variable with less levels: `USA`, `UK`, `Others`.

```{r}
levels(movie_metadata$country) <- c(levels(movie_metadata$country), "Others")
movie_metadata$country[(movie_metadata$country != 'USA')&(movie_metadata$country != 'UK')] <- 'Others' 
movie_metadata$country <- factor(movie_metadata$country)
table(movie_metadata$country)
```

Now that we've cleaned up our dataset, we can now continue to explore our data even further! In the next section we will be looking at genres, movie popularity, gross, profit, and many more other aspects pertinent to our data. 

-----

## Analysing Data

When inspecting a dataset of movies over the past few years, various interesting inferences can be uncovered. A movie may have a high rating yet low return on investment. Which genre is the most successful? Which actors are the most popular? These are some of the questions we aim to answer in this section.

We can start by performing basic analyis on our data. Thereafter, we delve a bit deeper into more specific parts of the dataset, in hopes of uncovering interesting observations.

### Basic Analysis

Let's first have a look at the number of movies that are produced over the years.

```{r fig.align="center"}
ggplot(movie_metadata, aes(title_year)) +
  geom_bar() +
  labs(x = "Year movie was released", y = "Movie Count", title = "Number of Movies Released Per Year (1916 - 2016)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(xintercept=c(1980), linetype="dotted")
```

From the graph, we see there aren’t many records of movies released before 1980. It’s better to remove those records because they might not be representative.

```{r}
movie_metadata <- movie_metadata[movie_metadata$title_year >= 1980,]
```

Let's have a look at the movie counts now:

```{r fig.align="center"}
ggplot(movie_metadata, aes(title_year)) +
  geom_bar() +
  labs(x = "Year movie was released", y = "Movie Count", title = "Number of Movies Released Per Year (1980 - 2016)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Movie Genre Analysis

```{r include=FALSE}
  
# setting up plotly label, axis and text customizations
f1 <- list(
  family = "Old Standard TT, serif",
  size = 14,
  color = "grey"
)
f2 <- list(
  family = "Old Standard TT, serif",
  size = 10,
  color = "black"
)
a <- list(
  titlefont = f1,
  showticklabels = T,
  tickangle = -45,
  tickfont = f2
)

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

# annotations for subplot
a1 <- list(x = 0.5, y = 1.0,
          showarrow = FALSE, 
          text = "Distribution of bugdet", 
          xanchor = "center", 
          xref = "paper", 
          yanchor = "bottom", 
          yref = "paper", 
          font = f1)
  
b1 <- list(x = 0.5, y = 1.0, 
           showarrow = FALSE, 
           text = "Distribution of gross", 
           xanchor = "center", 
           xref = "paper", 
           yanchor = "bottom", 
           yref = "paper",
           font = f1)
```

Now we can delve into more specific things regarding movies, like `genres`. 

#### Top Genres

```{r fig.align="center"}
genre = movie_metadata['genres']
genre = data.frame(table(genre))
genre = genre[order(genre$Freq,decreasing=TRUE),]

# Top 20 genres with the most movies
ggplot(genre[1:20,], aes(x=reorder(factor(genre), Freq), y=Freq, alpha=Freq)) + 
  geom_bar(stat = "identity", fill="blue") + 
  geom_text(aes(label=Freq),hjust=1.2, size=3.5)+
  xlab("Genre") + 
  ylab("Number of Movies") + 
  ggtitle("Top 20 genres with the most movies") + 
  coord_flip()
```

#### Split Genres

As you can see, movies have multiple genres that its associated with. For analysis purposes, we choose to use the first word in the genre column, as this is likely the most accurate description.

```{r}
head(movie_metadata$genres)
```

Let's split the genres separated by "|" into 8 different columns.

```{r fig.height=5, fig.width=11, message=FALSE, warning=FALSE}
# Split on "|"
genres_split <- str_split(movie_metadata$genres, pattern="[|]", n=2)

# Create Matrix
genres_matrix <- do.call(rbind, strsplit(movie_metadata$genres, '[|]'))

# Dataframe of genres
genres_df <- as.data.frame(genres_matrix)
```

`genre_df` consists of 8 columns, each with different genres. Let's have a look at the frequency of all the genres. 

```{r fig.align="center"}
# Collapse all genres into one column
genres_one_col <- gather(genres_df) %>% 
  select(value)

# Plot frequency of first column
genres_one_col %>%
  group_by(value) %>% 
  tally() %>% 
  filter(n >= 30) %>% 
  ggplot() +
  geom_bar(aes(x = value, y=n), stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Genre") +
  ylab("Frequency") +
  NULL
```
It is evident that the Drama and Comedy genre are the most popular to be produced. However, this does not mean that they are the most profitable, returning successful ROI’s. This will further be explored.


Previously we assumed that the first genre is the most applicable, therefore, we choose the first column as the genre for the movie and append it to the dataframe. 
```{r}
# Take old genres column away
movie_metadata <- subset(movie_metadata, select = -c(genres))

# Take first column of genres_df and add it to MAIN df
movie_metadata$genre <- genres_df$V1
```

How does this distribution look like over the years? Lets have a look at the frequency of genres between the period of 1980 and 2016.

```{r fig.align="center"}
movie_metadata %>%
 group_by(title_year, genre) %>%
 summarise(count = n()) %>%
 ggplot(aes(title_year, as.factor(genre))) +
 geom_tile(aes(fill=count),colour="white") +
 scale_fill_gradient(low="light blue",high = "dark blue") +
 xlab("Year of Movie") +
 ylab("Genre of Movie") +
 ggtitle("Heat Map of Movie Genres Produced Over the Years") +
  theme(panel.background = element_blank())
```

#### Which Genres are Popular?

In our dataset we have Facebook Likes and IMDB scores associated with a movie. This can help give us an indication of how popular each genre is. 

```{r include=FALSE}
# creating a function for plotting histogram using facet wrap
facet_hist_plot <- function(x, xlabel, bwidth){
  
  hp <- ggplot(data = movie_metadata, mapping = aes(x = x))
  hp + geom_histogram(aes(fill = genre), binwidth = bwidth,
                      show.legend = F, 
                      color = "black", size = 0.2, 
                      alpha = 0.8) +
    xlab(xlabel) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text = element_text(size = 12, angle = 20),
          axis.title = element_text(size = 14, 
                                    family = "Times", 
                                    color = "darkgrey",
                                    face = "bold")) + 
    facet_wrap(~ genre, scales = "free_y", ncol = 4)    
  }
```

```{r}
# Add column for total facebook likes
movie_metadata$total_facebook_likes <- movie_metadata$director_facebook_likes + 
  movie_metadata$actor_3_facebook_likes + movie_metadata$actor_1_facebook_likes +
  movie_metadata$cast_total_facebook_likes + movie_metadata$actor_2_facebook_likes + 
  movie_metadata$movie_facebook_likes
```

```{r include=FALSE}
movie_metadata <- movie_metadata[!is.na(movie_metadata$total_facebook_likes), ]
```

```{r, fig.height = 4, fig.width = 6, fig.align="center"}
# creating a data frame containing avg score, avg votes and avg fb likes
score_votes_likes <- movie_metadata %>% 
  group_by(genre) %>%
  summarise(count = n(),
            avg_score = round(mean(imdb_score), 1),
            avg_votes = mean(num_voted_users),
            avg_facebook_likes = mean(total_facebook_likes),
            avg_reviews = mean(num_user_for_reviews)) %>%
  filter(count > 10)

# arranging data frame by average score
arr_score <- arrange(score_votes_likes, desc(avg_facebook_likes))

score_votes_likes
```

##### Genres VS IMDB Scores

```{r fig.height = 10, fig.width = 10, fig.align="center"}  
# genre wise distribution of scores using a custom plot function  
facet_hist_plot(movie_metadata$imdb_score, 
                "Genre",
                bwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 9.3, 1)) 
``` 


```{r include=FALSE, fig.align="center"}
# creating a function for ploting bar graphs
bar_plot <- function(data, x, y, info, xlabl, ylabl, title, 
                      deci = NULL, suf = NULL){
  if(is.null(suf)) suf <- ""
  if(is.null(deci)) deci <- 0
  b1 <- ggplot(data, aes(x = reorder(genre, x), 
                               y = y, 
                               text = paste("Genre:", genre,
                                            info, 
                                            round(y, deci), suf)))
  
  b1 + geom_bar(aes(fill = genre), stat = "identity", 
           show.legend = F, color = "black", size = 0.2,
           width = 0.7, alpha = 0.7) +
  xlab(xlabl) +
  ylab(ylabl) +
  ggtitle(title) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 14, 
                                  color = "grey", family = "Times")) +
  scale_fill_brewer(palette = "Spectral")

ggplotly(tooltip = "text") %>%
  layout(margin = m, xaxis = a, yaxis = a)
}
```


```{r fig.align="center"}
# plotting 10 highest rated genres 
bar_plot(arr_score[1:10,],
         arr_score[1:10,]$avg_score,
         arr_score[1:10,]$avg_score,
         "| Average Score:",
         "Genres",
         "Average IMDB Score",
         "Top 10 IMDB Scores Per Genre",
         deci = 1)
```

##### Genres VS Facebook Likes

Let's now have a look at the Facebook likes. 

```{r fig.align="center"}

# plotting 10 highest rated genres 
bar_plot(arr_score[1:10,],
         arr_score[1:10,]$avg_facebook_likes,
         arr_score[1:10,]$avg_facebook_likes,
         "| Average Score:",
         "Genres",
         "Average Facebook Likes",
         "Top 10 Facebook Likes Per Genre",
         deci = 1)
```

### Popularity Analysis

IMDB ratings VS Movie Count

```{r fig.align="center"}
ggplot(movie_metadata, aes(imdb_score)) +
 geom_histogram(bins = 50) +
 geom_vline(xintercept = mean(movie_metadata$imdb_score,na.rm = TRUE),colour = "blue") +
 ylab("Movie Count") +
 xlab("IMDB Rating") +
 ggtitle("IMDB Ratings for Movies") +
  annotate("text", label = "Mean IMDB rating",x = 6.2, y = 50, size = 3, colour = "yellow",angle=90)
```

Popularity over the years
```{r message=FALSE, fig.align="center"}
#Creating the required subset of data 
movies_pop <- movie_metadata %>%
 select(title_year, movie_facebook_likes) %>%
 filter(title_year > 1980) %>%
 group_by(title_year) %>%
 summarise(avg = mean(movie_facebook_likes)) 

#Generating the popularity Vs time plot
 ggplot(movies_pop, aes(x = title_year, y = avg)) +
   geom_point() +
   geom_smooth() + 
   geom_vline(xintercept = c(1990,2004),colour = c("orange","blue")) +
   ylab("Mean Popularity Score") +
   xlab("Years") +
   annotate("text", label = "Facebook",x = 2003, y = 80, size = 3, colour = "blue",angle=90)+
   annotate("text", label = "IMDB",x = 1989, y = 80, size = 3, colour = "orange",angle=90)
```


Facebook Likes VS IMDB Score
```{r}
movie_metadata %>%
  plot_ly(x = ~movie_facebook_likes, y = ~imdb_score, color = ~content_rating , mode = "markers", text = ~content_rating, alpha = 0.7, type = "scatter")
```


Top 20 directors with highest average IMDB score
```{r}
movie_metadata %>%
  group_by(director_name) %>%
  summarise(avg_imdb = mean(imdb_score)) %>%
  arrange(desc(avg_imdb)) %>%
  top_n(20, avg_imdb) %>%
  formattable(list(avg_imdb = color_bar("orange")), align = 'l')
```

IMDB rating systen started in 1990’s. Social media platforms like Facebook had started in the mid 2000’s.
```{r}
#Performing operations on Movies Vote Count over the years
 movies_vote1 <- movie_metadata %>%
  select(title_year, num_voted_users) %>%
  group_by(title_year) %>%
  summarise(count = sum(num_voted_users)) 

 ggplot(movies_vote1, aes(x = title_year, y = count/1000)) +
   geom_bar( stat = "identity") +
   geom_vline(xintercept = c(1990,2004),colour = c("orange","blue")) +
   ylab("Vote count (in thousands)") +
   xlab("Years") +
   annotate("text", label = "Facebook",x = 2003, y = 160, size = 3, colour = "blue",angle=90) + 
   annotate("text", label = "IMDB",x = 1989, y = 160, size = 3, colour = "orange",angle=90)
```

Vote Counts VS IMDB score
```{r  fig.height = 4, fig.width = 6, fig.align="center"}
# plotting a boxplot for votes vs imdb_score  
bp <- na.omit(movie_metadata) %>% 
  ggplot(aes(x = vote_bucket, y = imdb_score)) +
  geom_boxplot(aes(fill = vote_bucket), alpha = 0.7,
               show.legend = F) +
  stat_summary(fun.y = mean, geom = "point",
               shape = 23) +
  coord_flip() +
  xlab("User Votes") +
  ylab("IMDB Score") +
  ggtitle("Box plot for User Votes vs IMDB Score") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, 
                                  color = "darkgrey",
                                  family = "Times"))
ggplotly(bp) %>%
  layout(margin = m, 
        xaxis = a, 
        yaxis = a,
        legend = list(orientation = "h", size = 4,
                       bgcolor = "#E2E2E2",
                       bordercolor = "darkgrey",
                       borderwidth = 1,
                       x = 0,
                       y = -0.3)) 
```

-----

## Network Analysis
