---
title: "An Exploratory Data and Network Analysis of Movies"
author: Clarice, Daven, Lucia, Christopher and Indurain
date: July 31, 2016
output:
  prettydoc::html_pretty:
    theme: tactile
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Resize Code ChunK Font Size
```{css, echo=FALSE}
code {
  font-size:0.8em;
}

table {
  font-size:0.9em;
}

p {
  font-size:0.9em;
}
```

## Introduction

In this report, we will be analysing a dataset from [Kaggle](https://www.google.com), which contains movies of different genres produced over a vast number of years. What makes this analysis interesting is that we can try and draw various conclusions based on a movie's popularity, directors or actors involved, year of production, and so forth. Moreover, we can construct various networks in an attempt to find meaningful and interesting results 

We have split our Exploratory Data Analysis into four main parts:

|  Section 	|                                                                                                                                                                                                                                                 	|
|---	|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------	|
| 1 	| Introducing the Data <br> - We first try to understand the data and look at its content.  |
| 2 	| Pre-Processing <br>- We look at what needs to be altered or removed from the dataset. <br>- We try clean any dirty text. <br>- We try to minimise the dataset's missing values.                                                                                    	|
| 3 	| Exploring the Data <br>- We conduct basic analysis on the dataset.  <br>- We explore genres. <br>- We explore movie popularity. <br>- We look at profit, gross, and return of interests with movies. <br>- We conduct more advanced analysis on the dataset.               	|
| 4 	| Network Analysis <br>- We measure the network (centrality, degree distribution, number of components, average degree) <br>- We use network measures to highlight certain nodes (actors) and see which measures of an actor will increase ratings and budgets.  	|                                                                                                                                  |

-----

## Admin

Before we start, let's keep this code chunk for importing the correct libraries and loading the appropriate dataset. We use pacman to load the following:

```{r}
pacman::p_load(rjson, jsonlite, DT,  RJSONIO, data.table, dplyr, compareDF, prettydoc, rmdformats)
```

We import the dataset like this:
```{r}
movie_metadata <- read.csv("../data/movie_metadata.csv", sep=";")
```

In the next section we can introduce our dataset and look its content. 

-----

## Introducing The Dataset

This section of the report is quite essential for our analysis. We cannot make any interesting inferences from the dataset if we do not know what is contained within it. In this section we will try to understand exactly what we are dealing with. Thereafter, we can begin to draw interesting results. We have already read in our dataset called `movie_metadata`, so we can see the following:

The dataset contains 28 unique columns/variables, each of which are described in the table below:

Variable Name            | Description
-------------------------|----------------------------------------------
color                    | Specifies whether a movie is black and white or color
director_name            | Contains name of the director of a movie
num_critic_for_reviews   | Contains number of critic reviews per movie
duration                 | Contains duration of a movie in minutes
director_facebook_likes  | Contains number of facebook likes for a director
actor_3_facebook_likes   | Contains number of facebook likes for actor 3
actor_2_name             | Contains name of 2nd leading actor of a movie
actor_1_facebook_likes   | Contains number of facebook likes for actor 1 
gross                    | Contains the amount a movie grossed in USD
genres                   | Contains the sub-genres to which a movie belongs
actor_1_name             | Contains name of the actor in lead role
movie_title              | Title of the Movie
num_voted_users          | Contains number of users votes for a movie
cast_total_facebook_likes| Contains number of facebook likes for the entire cast of a movie
actor_3_name             | Contains the name of the 3rd leading actor of a movie
facenumber_in_poster     | Contains number of actors faces on a movie poster
plot_keywords            | Contains key plot words associated with a movie
movie_imdb_link          | Contains the link to the imdb movie page
num_user_for_reviews     | Contains the number of user generated reviews per movie
language                 | Contains the language of a movie
country                  | Contains the name of the country in which a movie was made
content_rating           | Contains maturity rating of a movie
budget                   | Contains the amount of money spent in production per movie
title_year               | Contains the year in which a film was released
actor_2_facebook_likes   | Contains number of facebook likes for actor 2
imdb_score               | Contains user generated rating per movie
aspect_ratio             | Contains the size of the aspect ratio of a movie
movie_facebook_likes     | Number of likes of the movie on its Facebook Page


Furthermore, the dataset contains `5043` movies, spanning accross `96` years in 46 countries. There are `1693` unique director names and `X` number of actors/actresses. Around `X%` of the movies are from the USA, `X%` from UK, and `X%` from other countries.

-----

## Pre-Processing Data

When inspecting a database of films from recent years, various interesting inferences are uncovered. A film may have a high rating yet low return on investment. Which genre would you guess is the most successful? Which actors do you think are the most popular? 

-----

## Analysing Data

-----

## Network Analysis
