---
title: "Questions"
author: "Lucia"
date: "03/09/2019"
output: html_document
---

## Installed Libraries:

```{r}
library(tidyverse)
library(plotly)
library(RColorBrewer)
library(stringr)
library(DT)
```

## The dataset was imported through R:

```{r # Load data:I imported the dataset through R}
```


## Tidying and updating the dataset:   
```{r # fixing movie names}
movie_metadata$imdb_score <- gsub("Â", "", movie_metadata$imdb_score)
```


```{r # fixing movie names}
movie_metadata$imdb_score <- gsub("Â", "", movie_metadata$imdb_score)
```


```{r # extracting genres from sub genres}
movie_metadata$sub_genres <- movie_metadata$genres 
movie_metadata$genres <- NULL
movie_metadata$genres <- factor(str_extract(movie_metadata$sub_genres, pattern = "^[a-zA-Z]*"))
```


```{r # creating a new feature called total_facebook_likes}
movie_metadata$total_facebook_likes <- movie_metadata$director_facebook_likes + 
  movie_metadata$actor_3_facebook_likes + movie_metadata$actor_1_facebook_likes +
  movie_metadata$cast_total_facebook_likes + movie_metadata$actor_2_facebook_likes
```


```{r # converting imdb_score to numeric feature}
movie_metadata$imdb_score <- as.numeric(movie_metadata$imdb_score)
```


```{r # creating feature return_on_investment}
movie_metadata$budget <- movie_metadata$budget/1000000
movie_metadata$gross <- movie_metadata$gross/1000000
movie_metadata$return_on_investment <- (movie_metadata$gross - movie_metadata$budget)/movie_metadata$budget
movie_metadata$profit_loss <- ifelse(movie_metadata$return_on_investment > 0, 1, 0)  
```


## Checking how the new dataset looks and the new feature sets:

```{r # checking dimensions}
dimension <- dim(movie_metadata)
cat(dimension[1], "Items\n", dimension[2], "Fields")
cat("\n\n")
cat("Updated feature set:\n")

# checking categorical and quantitive features
var_type <- split(names(movie_metadata), sapply(movie_metadata, function(x) paste(class(x), collapse = " ")))
var_type  
```


## Structuring the dataset:

```{r # Structure}  
str(movie_metadata)  
```  


## Custom plot functions and plot styles for later uses:

```{r, echo=T}  
  
# setting up plotly label, axis and text customizations
f1 <- list(
  family = "Old Standard TT, serif",
  size = 14,
  color = "grey"
)
f2 <- list(
  family = "Old Standard TT, serif",
  size = 10,
  color = "black"
)
a <- list(
  titlefont = f1,
  showticklabels = T,
  tickangle = -45,
  tickfont = f2
)

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

# annotations for subplot
a1 <- list(x = 0.5, y = 1.0,
          showarrow = FALSE, 
          text = "Distribution of bugdet", 
          xanchor = "center", 
          xref = "paper", 
          yanchor = "bottom", 
          yref = "paper", 
          font = f1)
  
b1 <- list(x = 0.5, y = 1.0, 
           showarrow = FALSE, 
           text = "Distribution of gross", 
           xanchor = "center", 
           xref = "paper", 
           yanchor = "bottom", 
           yref = "paper",
           font = f1)
```


## Here are some custom plot functions and plot styles for later use:

```{r # Required graph setup}
# creating a function called scatter_plot for
# plotting scatter plots using ggplot and plotly

scatter_plot <- function(x, y, xlabel, ylabel, title,
                         text1, text2, text3,
                         alpha = NULL){
  if(is.null(alpha)) alpha <- 0.4
  gp <- ggplot(data = movie_metadata, mapping = aes(x = x, y = y,
                                          text = paste(text1, x,
                                                       text2, y,
                                                       text3, movie_title)))
  plot <- gp + geom_point(position = "jitter", 
                          show.legend = F, shape = 21, 
                          stroke = .2, alpha = alpha) + 
    xlab(xlabel) +
    ylab(ylabel) +
    ggtitle(title) +
    theme_minimal() +
    theme(legend.position = "none",
          plot.title = element_text(size = 12, face = "bold", 
                                    family = "Times", 
                                    color = "darkgrey")) 
  
  ggplotly(plot, tooltip = "text") %>%
    layout(m, xaxis = a, yaxis = a)
}

# creating function for plotting scatter plot using facet wrap
facet_plot <- function(x, y, xlabel, ylabel, alpha = NULL){
   if(is.null(alpha)) alpha <- 1
  
  fp <- ggplot(data = movie_metadata, mapping = aes(x = x, y = y))
  fp + geom_point(aes(fill = genres), position = "jitter", 
                  show.legend = F, shape = 21,
                  stroke = 0.2, alpha = alpha) +
    xlab(xlabel) +
    ylab(ylabel) +
    facet_wrap(~genres, scales = "free") +
    theme_minimal()
}

# creating a function for plotting a simple histogram
hist_plot <- function(x, xlabel, bwidth, fill = NULL, color = NULL){
  if(is.null(fill)) fill <- "orange"
  if(is.null(color)) color <- "black"
  hp <- ggplot(data = movie_metadata, mapping = aes(x = x))
  gp <- hp + geom_histogram(binwidth = bwidth, fill = fill, 
                            color = color, 
                            size = 0.2,
                            alpha = 0.7,
                            show.legend = F) +
    xlab(xlabel) +
    theme_minimal()
  
  ggplotly(gp) %>%
    layout(margin = m, xaxis = a, yaxis = a)
}

# creating a function for plotting histogram using facet wrap
facet_hist_plot <- function(x, xlabel, bwidth){
  
  hp <- ggplot(data = movie_metadata, mapping = aes(x = x))
  hp + geom_histogram(aes(fill = genres), binwidth = bwidth,
                      show.legend = F, 
                      color = "black", size = 0.2, 
                      alpha = 0.8) +
    xlab(xlabel) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text = element_text(size = 12, angle = 20),
          axis.title = element_text(size = 14, 
                                    family = "Times", 
                                    color = "darkgrey",
                                    face = "bold")) + 
    facet_wrap(~ genres, scales = "free_y", ncol = 4)    
  }


# creating function for plotting histograms for budget and gross
budg_gross_hist <- function(x){
  
  bh <- ggplot(movie_metadata, aes(x = x))
  bh + geom_histogram(binwidth = 0.05, 
                      fill = sample(brewer.pal(11, "Spectral"), 1),
                      color = "black", 
                      size = 0.09,
                      alpha = 0.7) + 
    scale_x_log10() + 
    theme_minimal()
  
  ggplotly() %>% 
  layout(m, xaxis = a, yaxis = a) 
}
                           

# creating a function for ploting bar graphs
bar_plot <- function(data, x, y, info, xlabl, ylabl, title, 
                      deci = NULL, suf = NULL){
  if(is.null(suf)) suf <- ""
  if(is.null(deci)) deci <- 0
  b1 <- ggplot(data, aes(x = reorder(genres, x), 
                               y = y, 
                               text = paste("Genre:", genres,
                                            info, 
                                            round(y, deci), suf)))
  
  b1 + geom_bar(aes(fill = genres), stat = "identity", 
           show.legend = F, color = "black", size = 0.2,
           width = 0.7, alpha = 0.7) +
  xlab(xlabl) +
  ylab(ylabl) +
  ggtitle(title) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 14, 
                                  color = "grey", family = "Times")) +
  scale_fill_brewer(palette = "Spectral")

ggplotly(tooltip = "text") %>%
  layout(margin = m, xaxis = a, yaxis = a)
}


# creating a function for plotting plotly line graph for title_year
line_graph <- function(data, y, name){
 
  scat_p1 <- plot_ly(data, x = ~title_year, y = ~ y, 
             name = name, type = 'scatter', mode = 'lines',
             line = list(color = sample(brewer.pal(11, "Spectral"), 1))) %>%
  layout(xaxis = list(title = "Title Year", zeroline = F,
                      showline = F, 
                      showticklabels = T),
         yaxis = list(title = "Average Score"),
         title = "Line Graph for Avg Score/Avg Votes/Avg User Review by Title Year", 
         font = list(family = "Serif", color = "grey"),
         legend = list(orientation = "h", size = 6,
                       bgcolor = "#E2E2E2",
                       bordercolor = "darkgrey",
                       borderwidth = 1),
         margin = m)
  scat_p1
  }
  
```  


## This is where the diagrams begin: ##

> Act 1: Early Exploration

1. Genre wise share of movies for top 10 genres [ ]

```{r fig.height = 4, fig.width = 6, fig.align = "center"}

# setting up color palette
colors <- c(brewer.pal(n = 11, name = "Spectral"))

# Plotting a donut chart by chaining commands together 
movie_metadata %>%
  group_by(genres) %>%
  summarise(count = n()) %>% 
  arrange(desc(count)) %>%
  head(10) %>%
  plot_ly(labels = ~genres, values = ~count,
          insidetextfont = list(color = 'Black'),
          marker = list(colors = colors, 
                        line = list(color = 'Black', width = .5)), 
          opacity = 0.8) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Genre wise share of movies for top 10 genres", 
         titlefont = list(family = "Times", size = 20, color = "grey"),
         xaxis = list(showgrid = T, zeroline = F, showticklabels = F),
         yaxis = list(showgrid = T, zeroline = F, showticklabels = F),
         showlegend = F, 
         margin = list(t = 50, b = 50))
```


2. Genre wise distribution of scores [ ]
> All genres with Imdb rating distribution

```{r fig.height = 10, fig.width = 10, fig.align="center"}  

# genre wise distribution of scores using a custom plot function  
facet_hist_plot(movie_metadata$imdb_score, 
                movie_metadata$genres,
                bwidth = 0.1)+
  scale_x_continuous(breaks = seq(0, 9.3, 1)) 
``` 



3. Bar chart for 10 highest rated genres [ ]
> Genres VS Average IMDB Score

```{r, fig.height = 4, fig.width = 6, fig.align="center"}

# creating a data frame containing avg score, avg votes and avg fb likes
score_votes_likes <- movie_metadata %>% 
  group_by(genres) %>%
  summarise(count = n(),
            avg_score = round(mean(imdb_score), 1),
            avg_votes = mean(num_voted_users),
            avg_facebook_likes = mean(total_facebook_likes),
            avg_reviews = mean(num_user_for_reviews)) %>%
  filter(count > 10)

# arranging data frame by average score
arr_score <- arrange(score_votes_likes, desc(avg_score))

# plotting 10 highest rated genres 
bar_plot(arr_score[1:10,],
         arr_score[1:10,]$avg_score,
         arr_score[1:10,]$avg_score,
         "| Average Score:",
         "Genres",
         "Average IMDB Score",
         "Bar chart for 10 highest rated genres",
         deci = 1)
```


4. Bar chart for 10 most voted genres [ ]
> Genres VS Average User Votes

```{r fig.height = 4, fig.width = 6, fig.align="center"}
# arranging data frame by avg votes
arr_votes <- arrange(score_votes_likes, desc(avg_votes))

# plotting bar chart for 10 most voted genres
bar_plot(arr_votes[1:10,],
         arr_votes[1:10,]$avg_votes,
         arr_votes[1:10,]$avg_votes,
         "| Average Votes:",
         "Genres",
         "Average User Votes",
         "Bar chart for 10 most voted genres")
```


5. Bar chart for 10 most FB liked genres [ ]
> Genres VS Average Facebook Likes

```{r fig.height = 4, fig.width = 6, fig.align="center"}
# arranging data frame by average fb likes
arr_fb_likes <- arrange(score_votes_likes, desc(avg_facebook_likes))

# plotting bar chart for top 10 genres by fb likes
bar_plot(arr_fb_likes[1:10,],
         arr_fb_likes[1:10,]$avg_facebook_likes,
         arr_fb_likes[1:10,]$avg_facebook_likes,
         "| Average FB Likes:",
         "Genres",
         "Average Facebook Likes",
         "Bar chart for 10 most FB liked genres")
```



6. Distribution of budget and gross [ ]

```{r  fig.height = 4, fig.width = 8, fig.align="center"}    
options(scipen = 999)  
# transformed budget histograms
p1 <- budg_gross_hist(movie_metadata$budget) %>% 
      layout(annotations = a1) %>%

# transformed gross histograms
p2 <- budg_gross_hist(movie_metadata$gross) %>%
  layout(annotations = b1) %>%

subplot(p1, p2, widths = c(0.5, 0.5))
```


7. Share of movies making profit or loss [ ]
> Did profitable movies recieve more appreciation by the users?

```{r fig.height = 4, fig.width = 6, fig.align="center"}

movie_metadata <- na.omit(movie_metadata)

# Plotting donut chart for share of movies making profit or loss 
movie_metadata %>%
  group_by(profit_loss) %>%
  summarise(count = n()) %>% 
  plot_ly(labels = ~factor(profit_loss), values = ~count,
          insidetextfont = list(color = 'Black'),
          marker = list(colors = sample(brewer.pal(11, "Spectral")), 
                        line = list(color = 'Black', width = .5)), 
          opacity = 0.8) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Share of movies making profit or loss", 
         titlefont = list(family = "Times", size = 20, color = "grey"),
         xaxis = list(showgrid = T, zeroline = F, showticklabels = F),
         yaxis = list(showgrid = T, zeroline = F, showticklabels = F),
         showlegend = T, 
         margin = list(t = 50, b = 50))

# Green is loss, Light Green is profit

```


> Act 2: Trends and Relationships

8. Line Graph for Avg Score/Avg Votes/Avg User Review by Title Year [ ]

```{r fig.height = 6, fig.width = 10, fig.align="center"}
# creating a data frame for average score, avg votes and, avg user reviews
# by title year
scat_year_score <- movie_metadata %>%
  group_by(title_year) %>%
  summarise(count = n(),
            IMDB_Score = round(mean(imdb_score),1),
            avg_votes = mean(num_voted_users),
            avg_user_review = round(mean(num_user_for_reviews)))

# plotting line graph for Avg score by title year 
lp1 <- line_graph(scat_year_score, 
                  scat_year_score$IMDB_Score, 
                  "Average Score")

# plotting line graph for Avg votes by title year
lp2 <- line_graph(scat_year_score, 
                  scat_year_score$avg_votes, 
                  "Average Votes")

# plotting line graph for Avg reviews by title year
lp3 <- line_graph(scat_year_score,
                  scat_year_score$avg_user_review,
                  "Average User Review") 


subplot(lp1, lp2, lp3, nrows = 3, heights = c(0.33, 0.33, 0.33))

```



9. Scatter plot for User Votes vs IMDB Score [ ]
> User Votes VS IMDB Score

```{r fig.height = 4, fig.width = 6, fig.align="center"}
# scatter plot for user votes vs imdb score
scatter_plot(movie_metadata$num_voted_users, movie_metadata$imdb_score, 
     "User Votes",
     "IMDB Score",
     "Scatter plot for User Votes vs IMDB Score",
     "User Votes:",
     "<br>IMDB Score:",
     "<br>Title:",
     alpha = 0.3)
```



10. Box plot for User Votes vs IMDB Score
> IMDB Score VS User Votes

```{r echo=T}
# breaking num_users_voted into 4 buckets 
movie_metadata$vote_bucket <- cut(movie_metadata$num_voted_users, 
                         c(0, 50000, 100000, 300000, 500000))
```

```{r  fig.height = 4, fig.width = 6, fig.align="center"}
# plotting a boxplot for votes vs imdb_score  
bp <- na.omit(movie_metadata) %>% 
  ggplot(aes(x = vote_bucket, y = imdb_score)) +
  geom_boxplot(aes(fill = vote_bucket), alpha = 0.7,
               show.legend = F) +
  stat_summary(fun.y = mean, geom = "point",
               shape = 23) +
  coord_flip() +
  xlab("User Votes") +
  ylab("IMDB Score") +
  ggtitle("Box plot for User Votes vs IMDB Score") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, 
                                  color = "darkgrey",
                                  family = "Times"))
ggplotly(bp) %>%
  layout(margin = m, 
        xaxis = a, 
        yaxis = a,
        legend = list(orientation = "h", size = 4,
                       bgcolor = "#E2E2E2",
                       bordercolor = "darkgrey",
                       borderwidth = 1,
                       x = 0,
                       y = -0.3)) 
```


> Act 3: The Last Enquiry

_____
