---
title: "Questions"
author: "Luciaaaa"
date: "03/09/2019"
output: html_document
---

```{r setup, include=FALSE}
library(pacman)
pacman::p_load(dplyr, ggplot2, tidyr )
knitr::opts_chunk$set(echo = TRUE)
setwd(getwd())
```


```{r}
movie_metadata <- read.csv("../movie_metadata.csv", sep=";")
```

## Questions

Here is a list of questions we could play around with. Each title has "[ ]" on the side. If the question has been completed successfully, add an "X" inbetween the "[ ]", like "[X]". These questions have been arranged to URL links. Hopefully the source codes helps...

Also, read through ALLLLLL of the links. Let's start somewhere simple and work ourselves up. 

-----

### Link 1: https://rstudio-pubs-static.s3.amazonaws.com/384914_7bee8325659440ab9db1b3de36c37343.html

1. HeatMap [ ]
> Genre VS Year that Movie is Released

```{r}
movie_metadata %>%
 filter(title_year > 2000) %>%  #Filter movies released after 2000
 group_by(title_year, genres) %>%
 summarise(count = n()) %>% #Count of Movies released
 ggplot(aes(title_year, as.factor(genres)))+
 geom_tile(aes(fill=count),colour="white")+
 scale_fill_gradient(low="light blue",high = "dark blue") +
 xlab("Year of Movie release")+
 ylab("Genre of Movie Release")+
 ggtitle("Heat Map")+
theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

2. Movie Budgets [ ]

```{r}

```

3. Movie Gross [ ]

4. 3D Plot [ ]
> IMDb Rating VS Budget VS Gross

5. Vote counts over the years [ ]

6. Popularity over the years [ ]

7. Distribution of Imdb Ratings
> Movie Count VS Imdb Ratings

-----

### Link 2: https://rpubs.com/jasmine36/234304

1. Most Successfull Directors [ ]
> Director vs FB likes
> Director VS Gross
> Director VS Imdb
> Average Profit VS Number of Films (Points being Director)


2. Commercial Success VS Critical Acclaim [ ]
> Gross Money VS Imdb score (Profit being size of point; Point Colour being rating)
> Also available on: http://rstudio-pubs-static.s3.amazonaws.com/342210_7c8d57cfdd784cf58dc077d3eb7a2ca3.html

3. Top 20 genres with the most movies [ ]
> Genre VS Freq of movies
> Look at Link 4 for more inspo

-----

### Link 3: http://rstudio-pubs-static.s3.amazonaws.com/342210_7c8d57cfdd784cf58dc077d3eb7a2ca3.html
Literally works on our data, however, this link does a lot of pre-processing that we might need to take a look at. For example, they use the first word in the genre column as the genre of the movie. 

1. Number of Movies Produced Per Year [ ]
> Count of movies VS Year

2. Most Successfull Directors [ ]
> Table representation of Director and their average IMDb rating

3. Relationship between number of FB likes and Imdb score [ ]
> Imdb score VS movie FB likes

4. Correlation Heatmap [ ]

5. Full-Grown Tree [ ]
> Random Forest
> K-Nearest Neighbour

-----

#### Link 4: https://www.kaggle.com/karrrimba/understanding-movies-through-data
Literally works on our data. A lot of code to look trough. 

1. Genre wise share of movies for top 10 genres [ ]

2. Genre wise distribution of scores [ ]
> All genres with Imdb rating distribution

3. Bar chart for 10 highest rated genres [ ]
> Genres VS Average IMDB Score

4. Bar chart for 10 most voted genres [ ]
> Genres VS Average User Votes

5. Bar chart for 10 most FB liked genres [ ]
> Genres VS Average Facebook Likes

6. Distribution of budget and gross [ ]

7. Share of movies making profit or loss [ ]
> Did profitable movies recieve more appreciation by the users?

8. Line Graph for Avg Score/Avg Votes/Avg User Review by Title Year [ ]

9. Scatter plot for User Votes vs IMDB Score [ ]
> User Votes VS IMDB Score

10. Box plot for User Votes vs IMDB Score
> IMDB Score VS User Votes

_____


### Other Analysis Questions

1. Rating VS Budget [ ]

2. Rating VS Gross [ ]

3. Budget VS Gross [ ]

4. Most Successful/Unsuccessful Movies
> Based on their ratings

5. Highest Rated Actor
> Movie VS Imdb Score
> Most Popular Actor: Based on FB likes

-----


